\documentclass[12pt]{report}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\graphicspath{{images/}}
\usepackage{multirow}
\usepackage{float}
\usepackage{subcaption}
\usepackage[a4paper, width=150mm, top=25mm, bottom=25mm]{geometry}
\usepackage{fancyhdr}
\usepackage{amsmath}
\usepackage{titlesec}
\usepackage{comment}
\usepackage{pythonhighlight}
\usepackage[numbers]{natbib}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=black,
    filecolor=magenta,      
    urlcolor=blue,
    citecolor=black,
    pdftitle={Overleaf Example},
    pdfpagemode=FullScreen,
    }
\usepackage{mathtools}
\usepackage{tcolorbox}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{calc}
\newcommand\Myperm[2][^n]{\prescript{#1\mkern-2.5mu}{}P_{#2}}
\newcommand\Mycomb[2][^L]{\prescript{#1\mkern-0.5mu}{}C_{#2}}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

\titleformat{\chapter}[hang] 
{\normalfont\huge\bfseries}{\chaptertitlename\ \thechapter:}{1em}{}
\pagestyle{fancy}
\fancyhead{}
\fancyhead[L]{\thesection}
\fancyhead[R]{\textbf{Entropy using cellular automata}}
\fancyfoot{}
\fancyfoot[L]{\thepage}
\fancyfoot[R]{Bowen Shiro Husin}
\setlength\parindent{0pt}

\begin{document}
\input{titlepage.tex}

\chapter*{Acknowledgements}
This project is intended to keep my brain busy while I am 
\emph{job-searching}. Likewise, it's meant to be a coding challenge that I've come up with. This document is meant to declare the mathematics and my line of thinking in designing the rules and entropy computation. Should you read this part, feel free to use the code and tinker with it. If you are a physics teacher, feel free to use it to teach your students about the fascinating concept of entropy.   \par

\vspace{0.3cm}
To view how my code is implemented check my GitHub \href{https://github.com/ShiroHusin/Entropy_Simulation}{repository}.


\chapter*{Abstract}
Entropy is a fascinating concept with far-reaching consequences for our daily lives. In this document, a cellular-automata-inspired approach is proposed. Namely, a 2-dimensional array consisting of cells with integer values ranging from 0-8 is constructed for visualizing  entropy increase over time. \par

\vspace{0.3cm}
\noindent
For version 2, the Boltzmann factor alongside temperature is implemented to count for energy differences between cells and alter their probabilities. Likewise, a new method is used to compute the entropy of the simulation. Namely, the use of the bitwise XOR function is abandoned for version 2 and is replaced with a more elementary operation that is closer to how entropy is computed by the famous Boltzmann equation. Finally, a custom function is used to map the conductivity or $c$ with move probability or $\alpha$. \par

\vspace{0.3cm}
\noindent
Results from the simulation show that the underlying structure of the 2-dimensional array will become more uniformly distributed no matter the starting configuration. Plots show that both conductivity and temperature have tremendous effects on the final entropy after a set amount of iterations. Conductivity increases it asymptotically while temperature also flattens it asymptotically.  \par



\tableofcontents

\pagebreak

\include{chapters/Chapter1.tex}


\include{chapters/Chapter2.tex}


\include{chapters/Chapter3.tex}

\pagebreak
\include{chapters/Remarks.tex}

\end{document}
