\chapter{Closing remarks}
Overall, this was a passion project that is of my own doing. Well, not exactly. To be fair, 30\% of the code was done by Chat-GPT3. However, the problem-solving strategy and creative output are almost entirely my own work. In early January of this year, I had a realization. The fact is that I can use the principles of cellular automata and some probability in order to demonstrate why the 2nd law of thermodynamics works and that there is no mysterious force behind it. It is only the case that entropy always increases because it's much more likely to do so than not. \par

\vspace{0.3cm}
\noindent
Overall, this was a self-confidence boost that I desperately needed. After a very unproductive and depressing 2nd half of 2022, a part of me wanted to prove to myself that I could actually engage in high-level problem-solving and that I am capable of doing something intellectually demanding. Likewise, working on this was also an exorcism of my own demons of self-doubt and self-hatred that would otherwise bring me down to absolute mediocrity and waste the scientific skills and abilities that I acquired over the 22 years of my life. \par

\vspace{0,3cm}
\noindent
This was a work born out of desperation and I could safely say that I feel more accomplished than getting my testamur 3 months ago in December 2022! Anyways, life goes on and the show must go on. Life was a long hard journey but I decided should no longer betray what I wanted to do as a child, which is to be a scientist. 